---
title       : Predict your car's gas mileage in 1974
subtitle    : How should you trade off car features against budget after the oil crisis?
author      : David Modjeska
job         : UX Lead at PriceMetrix in Toronto
framework   : html5slides   # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
--- 

## Pitch: Predict your car's gas mileage in 1974  

#### How can you trade off features against economy after the oil crisis?  

###### David Modjeska  

---

## The Application

* Imagine that it's 1974. The oil crisis of 1973 has just happened. 
What kind of car should you buy?

* Meanwhile, you've heard a lot about linear regression. What's it useful for?
How can you see a demonstration?

* In a perfect world, you could use a computer to predict a car's gas mileage,
using linear regression to calculate the answer. Well, it's possible.

* There's an application called Mileage Predictor on the Web. 
With this app, you just enter basic information about a car.
Then you immediately get the predicted mileage.

* The application includes extra information for learning more --
a detailed trend graph, and background on the application and its statistics.

--- 

## An Example

* Let's consider an example.

* Imagine that you're interested in a car with 6 cylinders, a weight of 3.5 tons, and automatic transmission.

* What gas mileage will this type of car have? 

* A sample screen from the Mileage Predictor shows us the answer.

---

## The Prediction

```{r echo = FALSE, warning = FALSE }

library(datasets)
library(dplyr)

# read and clean data
data(mtcars)
mtcars1 <- mtcars %>%
    rename(Transmission = am)  %>%
    mutate(Transmission = gsub("0", "Automatic", Transmission)) %>%
    mutate(Transmission = gsub("1", "Manual", Transmission)) %>%
    mutate(Transmission = factor(Transmission)) %>%
    mutate(cyl = paste(cyl, " Cylinder", sep = "")) %>%
    mutate(cyl = factor(cyl)) %>%
    select(Transmission, cyl, wt, mpg)

# fit model - multiple linear regression
fit <-
    lm(data = mtcars1, mpg ~ cyl + wt + Transmission + wt:Transmission - 1)

    # define variable to hold sample input
    sample_data <- data.frame(cyl = "6 Cylinder", wt = 3.5, Transmission = "Automatic")

    # define variable to hold prediction
    predict_mpg <- round(predict(fit, newdata = sample_data)[1])

    predict_text <- paste("The predicted gas mileage of your car is ", round(predict_mpg), " MPG.")
```

 "The predicted gas mileage of your car is **`r predict_mpg`** MPG."

```{r echo = FALSE, warning = FALSE }   

library(grid)
library(ggplot2)

# define function to allocate graph colors algorithmically
# from http://stackoverflow.com/questions/8197559/emulate-ggplot2-default-color-palette
ggplot_colors <- function(n = 6, h = c(0, 360) + 15) {
    if ((diff(h) %% 360) < 1)
        h[2] <- h[2] - (360 / n)
    hcl(h = (seq(h[1], h[2], length = n)), c = 100, l = 65)
}

# create data frame to hold prediction and sample input
new_data <- data.frame(
    cyl = sample_data$cyl,
    wt = sample_data$wt,
    Transmission = sample_data$Transmission,
    mpg = predict_mpg
)

# create plot
coeff <- fit$coeff
plot_colors <- ggplot_colors(3)
ggplot(mtcars1, aes(x = wt, y = mpg, color = Transmission)) +
    geom_blank() +
    geom_abline(data = subset(mtcars1, 
                              cyl == '4 Cylinder' & Transmission == 'Manual'),
                aes(color = Transmission), size = 1.5,
                intercept = coeff[1], slope = coeff[4]
                ) +
    geom_abline(data = subset(mtcars1, 
                              cyl == '4 Cylinder' & Transmission == 'Automatic'),
                aes(color = Transmission), size = 1.5,
                intercept = coeff[1] + coeff[5],
                slope = coeff[4] + coeff[6]
                ) +
    geom_abline(data = subset(mtcars1, 
                              cyl == '6 Cylinder' & Transmission == 'Manual'),
                aes(color = Transmission), size = 1.5,
                intercept = coeff[2], slope = coeff[4]
                ) +
    geom_abline(data = subset(mtcars1, 
                              cyl == '6 Cylinder' & Transmission == 'Automatic'),
                aes(color = Transmission), size = 1.5,
                intercept = coeff[2] + coeff[5],
                slope = coeff[4] + coeff[6]
                ) +
    geom_abline(data = subset(mtcars1, cyl == '8 Cylinder' &
                                  Transmission == 'Manual'),
                aes(color = Transmission), size = 1.5,
                intercept = coeff[3], slope = coeff[4]
                ) +
    geom_abline(data = subset(mtcars1, cyl == '8 Cylinder' &
                                  Transmission == 'Automatic'),
                aes(color = Transmission), size = 1.5,
                intercept = coeff[3] + coeff[5],
                slope = coeff[4] + coeff[6]
                ) +
    facet_wrap( ~ cyl, ncol = 3) +
    # the following point is a dummy that causes the legend to display
    geom_point(data = subset(mtcars1,
                             cyl == "4 Cylinder" &
                                 Transmission == "Automatic" &
                                 wt == 1.5 &
                                 mpg == 23)
               ) +
    geom_point(data = new_data, pch = 21,
               fill = plot_colors[1], color = "white", size = 8
               ) +
    ggtitle("Your Car and General Mileage Trends\n") +
    xlab("Weight (Tons)") +
    ylab("Mileage (MPG)") +
    theme(axis.title.x = element_text(vjust = -0.5)) +
    theme(axis.title.y = element_text(vjust = 1.5)) +
    theme(legend.position = "bottom", legend.margin = unit(1.5, "lines")) +
    scale_color_manual(
        values = plot_colors[2:3],
        breaks = mtcars1$Transmission,
        labels = mtcars1$Transmission,
        name = "Transmission"
    )

```

--- 

## Try it yourself

* To try out the app yourself, please visit shinyapps.io:

     + https://dmodjeska.shinyapps.io/Mileage_Predictor/

* For technical details, the source code for this app can be found on GitHub:

     + https://github.com/dmodjeska/mileage-predictor
     
* Thank you for your time!




